Semi-Supervised Factored Logistic Regression for High-Dimensional Neuroimaging Data

### 2015-10-04 paper_nips/paper.tex
CAMERA READY

\author{Danilo Bzdok, Michael Eickenberg, Olivier Grisel,
  Bertrand Thirion,
  Ga\"el Varoquaux \\\textbf{\textit{email:} }firstname.lastname@inria.fr}

@@

To \textit{test}, on the other hand,
the neural correlates underlying mental tasks,
the general linear model (GLM) is the dominant approach \cite{friston94}.
The contribution of
individual brain voxels is estimated
according to a design matrix of experimental tasks.
Alternatively, psychophysiological interactions (PPI)
elucidate the influence of one brain region on another conditioned
by experimental tasks \cite{friston97}.
As a last example, an increasing number of
neuroimaging studies model
experimental tasks by training classification algorithms on brain signals
\cite{poldrack09decoding}.
All these methods are applied to task-associated (i.e., \textit{labeled})
data that capture brain dynamics
during stimulus-guided behavior.
Two important conclusions can be drawn.
First, the mentioned supervised neuroimaging analyses typically yield
results in a voxel space.
This ignores the fact that the BOLD
signal exhibits spatially distributed patterns
of coherent neural activity.
%
Second, existing supervised neuroimaging analyses
cannot exploit the abundance
of easily acquired resting-state data \cite{biswaldiscovery}.
These may allow better discovery of
the manifold of brain states
due to the high task-rest similarities of neural activity patterns,
as observed using ICA \cite{smith2009}.
% XXX: you don't discuss the absence of information sharing across tasks / cognitions, i.e. the multi-task aspect of your work.

Both these neurobiological properties can be conjointly exploited in an
approach that is
\textit{mixed} (i.e., using rest and task data),
\textit{factored} (i.e., performing network decomposition), and
\textit{multi-task} (i.e., capitalize on neural representations shared across
mental operations).
%
The integration of brain-network discovery into
supervised classification can yield a semi-supervised learning
framework. The most relevant
neurobiological structure should hence be identified
for the prediction problem at hand.
%
Autoencoders suggest themselves because they can emulate
variants of most unsupervised learning algorithms,
including PCA, SPCA, and ICA \cite{hinton06}.
\begin{wrapfigure}{r}{0.40\textwidth}
  \centering
  %\captionsetup{labelformat=nonumber}
    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
  \vspace{-0.7cm}
  \caption {\textbf{Model architecture}
  Linear autoencoders find an optimized compression of 79,941 brain voxels
  into $n$ unknown activity patterns by improving reconstruction from them.
  The decomposition matrix equates with the bottleneck of
  a factored logistic regression.
  Supervised multi-class learning on task data ($X_{task}$)
  can thus be guided by
  unsupervised decomposition of rest data ($X_{rest}$).
  }
\end{wrapfigure}
%
Autoencoders (AE)
are layered learning models
that condense the input data to
local and global 
representations
via reconstruction under compression prior.
%
% XXX Michael: In fact, they behave like a non-centered PCA, 
% XXX which amounts to just doing an SVD
% XXX which btw is NOT a convex optimization problem, although easy to 
% XXX calculate and with a global optimum.
% XXX Autoencoders are also not one-layered by default. Let's make sure that
% XXX we don't trip up by getting the technicalities wrong and getting 
% XXX reviewers annoyed at us because of it.
% DanilO: cool thanks. That is the stuff, we need.
They behave like a (truncated) PCA
in case of one linear hidden layer and a squared error loss
\cite{baldi1989neural}.
% This architecture yields an optimization objective
% with unique global minimum.
Autoencoders behave like a SPCA if shrinkage terms are added to the
model weights in the optimization objective.
Moreover, they have the characteristics of an ICA in case of tied weights
and adding a nonlinear convex function
at the first layer \cite{le2011ica}.
% XXX What convex transformation? Why is that relevant? Convexity is only
% XXX relevant for the loss
% Danilo: it's about adding a neural-network-like nonconvex function for
% first-layer activation
These authors further demonstrated that ICA, sparse autoencoders, and 
sparse coding are mathematically equivalent
under mild conditions.
%
Thus, autoencoders may flexibly project the neuroimaging data
onto the main directions of variation.

In the present investigation,
a linear autoencoder will be fit to
(unlabeled) rest data and
integrated as a rank-reducing bottleneck
into a multinomial
logistic regression fit to
(labeled) task data.
We can then solve the compound statistical problem of
unsupervised data representation and
supervised classification, previously studied in isolation.
%
From the perspective of dictionary learning, the first layer represents
% XXX Michael: This is the crucial difference: The first layer is an analysis
% XXX operation: it extracts information in some way. It may be extracting
% XXX activations for basis functions, in which case it would be the pseudo-
% XXX inverse of those basis functions. But unless they are orthogonal, these
% XXX weights definitely don't represent basis functions themselves. It is 
% XXX the second layer of the autoencoder that represents basis functions. If
% XXX weights are tied, they are one and the same thing, with interesting
% XXX implications. But in general, only the second layer represents a basis.
projectors to the discovered set of basis functions
which are linearly combined 
by the second layer to perform predictions \cite{olshausen96}.
% XXX Michael: ^ This needs to be changed around: This first layer obtains
% XXX activations, the second layer weights are basis functions which are
% XXX multiplied to these activations
%
Neurobiologically, this allows 
delineating a low-dimensional manifold
of brain network patterns and then 
distinguishing mental tasks
by their most discriminative linear combinations.
%
Theoretically, a reduction in model variance should be achieved by
resting-state autoencoders that
privilege the most neurobiologically
valid models in the hypothesis set.
%
Practically, neuroimaging research frequently suffers from
data scarcity. This limits the set of representations that can be
extracted from GLM analyses based on few participants.
%
We therefore contribute a computational framework that
1) analyzes many problems simultaneously
(thus finds shared representations by ``multi-task learning'')
and
2) exploits unlabeled data,
as they span a space of meaningful configurations.

\section{Methods}
%
\paragraph{Data.}
As the currently biggest openly-accessible reference dataset,
we chose resources from the Human Connectome Project (HCP)
\cite{barch2013}.
Neuroimaging task data with labels of ongoing cognitive processes
were drawn from 500 unrelated,
healthy HCP participants (cf. Appendix for details on datasets).
18 HCP tasks 
were selected that are known to elicit reliable neural activity
across participants (Table \ref{table_tasks}).
In sum, the HCP task data incorporated 8650 first-level activity maps
from 18 diverse paradigms administered to 498 participants (2 removed
due to incomplete data).
All maps were resampled to a common $60\times72\times60$ space of
3mm isotropic voxels and gray-matter masked (at least 10\% tissue
probability).
The supervised analyses were thus based on labeled HCP task maps with
79,941 voxels of interest representing z-values in gray matter.

@@

where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79{,}941$
voxel values from each
rest map,
$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$-dimensional hidden state
(i.e., distributed neural activity patterns), and 
$\mathbf{x' \in \mathbb{R}^{d}}$
is the reconstruction vector of the original activity map
from the hidden variables. 
% XXX use bold letters for vectors and matrices
Further, $\mathbf{W_0}$ denotes the weight matrix that
transforms
from input space into the hidden space (encoder),
$\mathbf{W_1}$ is the weight matrix for back-projection
from the hidden variables to the
output space (decoder).
$\mathbf{b_0}$ and $\mathbf{b_1}$ are corresponding bias vectors.
The model parameters $\mathbf{W_0, b_0, b_1}$ are found by
minimizing the expected squared reconstruction error
\begin{equation}
  \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
  \mathbb E\left[\| \mathbf{x} - \mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right].
\end{equation}

@@

\paragraph{Factored logistic regression.}
Our factored logistic regression model is best described as a
variant of a multinomial logistic regression.
Specifically, the weight
matrix is replaced by the product of two weight matrices with a common
latent dimension.
The later is typically much lower than the dimension of the data.
%
Alternatively, this model can be viewed as a single-hidden-layer
feed-forward neural network with a linear activation function for the
hidden layer and a softmax function on the output layer.
As the
dimension of the hidden layer is much lower than the input layer, this
architecture is sometimes referred to as a ``linear bottleneck'' in the
literature.
%
The probability of an input $\mathbf{x}$ to belong
to a class $i \in \{1, \dots, l\}$ is given by
\begin{equation}
  \begin{split}
    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_\mathcal{LR}(\mathbf x)),
  \end{split}
  \label{eq:lr}
\end{equation}
where \(f_\mathcal{LR}(\mathbf x) = \mathbf{V_1 (V_0 x + c_0) + c_1}\) computes 
multinomial logits and \(\softmax_i(x) = \exp(x_i)/\sum_j\exp(x_j)\).
The matrix $\mathbf{V_0 \in \mathbb{R}^{dxn}}$
transforms the input $\mathbf{x \in \mathbb{R}^{d}}$
into $n$ latent components
and the matrix $\mathbf{V_1 \in \mathbb{R}^{nxl}}$
projects the latent components
onto hyperplanes that reflect $l$ label probabilities.
$\mathbf{c_0}$ and $\mathbf{c_1}$ are
bias vectors.
The loss function is given by
\begin{equation}
    \mathbb E\left[{\mathcal{L_{LR}}}(\mathbf{x, y})\right] \approx %
\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1}).
\label{eq:lr_loss}
\end{equation}

\paragraph{Layer combination.}
Importantly, the optimization problem of the linear autoencoder
and the factored logistic regression
are linked in two ways. First, their transformation matrices mapping from
input to the latent space are tied
\begin{eqnarray}
  \mathbf{V_0} = \mathbf{W_0}.
\end{eqnarray}
We hence search for a compression of the 79,941 voxel values into $n$ unknown
components that represent a latent code optimized for both
rest and task activity data.
Second, the objectives of the autoencoder and the factored
logistic regression are interpolated in the common loss function
% XXX Michael: In resolving conflict I opted for my change. Feel free to revert if that was intended.
\begin{equation}
{\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
+ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \Omega(\theta).
  \label{eq:loss_equ}
\end{equation}

In so doing, we search for the combined model parameters
$\theta=\{\mathbf{V_0,V_1,c_0, c_1, b_0, b_1}\}$
with respect to the (unsupervised) reconstruction error and the
(supervised) task detection.
${\mathcal{L_{AE}}}$ is devided by ${N_{X_{rest}}}$ to equilibrate both
loss terms to the same order of magnitude.
\(\Omega(\theta)\) represents an ElasticNet-type regularization
that combines $\ell_1$ and $\ell_2$ penalty terms.
%$\forall p \in \theta$.

@@

\paragraph{Implementation.}
The analyses were performed in Python.
We used \textit{nilearn} to handle
the large quantities of neuroimaging data 
\cite{abrah14}
and
\textit{Theano} for automatic, numerically stable
differentiation of symbolic computation graphs
\cite{bastien2012theano, bergstra2010theano}.
All Python scripts that generated the results are
accessible online for reproducibility and reuse
(\url{http://github.com/anonymous/anonymous}).
% (url{http://github.com/banilo/nips2015}).

@@

 \multicolumn{1}{l|}{5}   & 45,1 \%        & 32,2 \%        & 37,5 \%       & 44,2 \%        & \textbf{95,7\%}      \\
 \multicolumn{1}{l|}{20}  & 78,1 \%        & 78,2 \%        & 81,0 \%       & 63,2 \%        & \textbf{97,3\%}      \\
 \multicolumn{1}{l|}{50}  & 81,7 \%        & 84,0 \%        & 84,2 \%       & 77,0 \%        & \textbf{97,6\%}      \\
 \multicolumn{1}{l|}{100} & 81,3 \%        & 82,2 \%        & 87,3 \%       & 76,6 \%        & \textbf{97,4\%}      \\ \hline

@@

Additionally, the ``out-of-study performance'' of

@@

\begin{table}[h]
  \centering
  \resizebox{1.00\textwidth}{!}{%
\tabcolsep=0.05cm
\begin{tabular}{l|ccccc|ccccc|ccccc}
  \hline
                            & \multicolumn{5}{c|}{\textit{n = 5}}                                                                                                                                                 & \multicolumn{5}{c|}{\textit{n = 20}}                                                                                                                                                      & \multicolumn{5}{c}{\textit{n = 100}}                                                                           \\ \hline
  \textit{}                 & \multicolumn{1}{l}{$\lambda=0$}    & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0$}  & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0$      & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
  Out-of-sample\\accuracy   & \textit{6,0\%}                     & 88,9\%                             & 95,1\%                            & \textbf{96,5\%}                    & 95,7\%                           & \textit{5,5\%}                   & 97,4\%                               & \textbf{97,8\%}                     & 97,3\%                               & 97,3\%                             & \textit{6,1\%}   & 97,2\%           & 97,0\%           & \textbf{97,8\%}  & 97,4\%                             \\
  Precision (mean)          & \textit{5,9\%}                     & 87,0\%                             & 94,9\%                            & \textbf{96,3\%}                    & 95,4\%                           & \textit{5,1\%}                   & \textbf{97,4\%}                      & 97,1\%                              & 97,0\%                               & 97,0\%                             & \textit{5,9\%}   & 96,9\%           & 96,5\%           & \textbf{97,5\%}  & 96,9\%                             \\
  Recall (mean)             & \textit{5,6\%}                     & 88,3\%                             & 95,2\%                            & \textbf{96,6\%}                    & 95,7\%                           & \textit{4,6\%}                   & \textbf{97,5\%}                      & 97,5\%                              & 97,4\%                               & 97,4\%                             & \textit{7,2\%}   & 97,2\%           & 97,2\%           & \textbf{97,9\%}  & 97,4\%                             \\
  F1 score (mean)           & \textit{4,1\%}                     & 86,6\%                             & 94,9\%                            & \textbf{96,4\%}                    & 95,4\%                           & \textit{3,8\%}                   & \textbf{97,4\%}                      & 97,2\%                              & 97,1\%                               & 97,1\%                             & \textit{5,3\%}   & 97,0\%           & 96,7\%           & \textbf{97,7\%}  & 97,2\%                             \\ \hline
  Out-of-study\\accuracy    & \textit{39,4\%}                    & 60,8\%                             & 54,3\%                            & 60,7\%                             & \textbf{62,9\%}                  & \textit{77,0\%}                  & 79,7\%                               & \textbf{81,9\%}                     & 79,7\%                               & 79,4\%                             & 79,2\%           & \textbf{82,2\%}  & 81,7\%           & 81,3\%           & \textit{75,8\%}                    \\ \hline
  \end{tabular}
}
\vspace{-0.2cm}
 \caption{\textbf{Performance of SSFLogReg across model parameter choices.}
 Chance is at 5,6\%.
 }
  \label{table_two}
\end{table}

We made three noteworthy observations (Table \ref{table_two}).
%
First, the most supervised estimator ($\lambda=1$) achieved in no
instance the best accuracy, precision, recall, or f1 scores on HCP data.
Classification by SSFLogReg is therefore facilitated by
imposing structure from the unlabeled rest data.
%
Second, the higher the number of latent components $n$,
the higher the out-of-study performance with small values of $\lambda$.
This suggests that the presence of more rest-data-inspired hidden components
results in more effective feature representations in unrelated task data.
%
Third, for $n=20$ and $100$ (but not $5$) the purely rest-data-trained
decomposition matrix ($\lambda=0$) resulted in
noninferior out-of-study performance
of 77,0\% and 79,2\%,
respectively (Table \ref{table_two}).
This confirms that guiding model learning by task-unrelated structure
extracts features of general relevance
beyond the supervised problem at hand.

@@

>>>>

\author{Danilo Bzdok, Michael Eickenberg, Olivier Grisel,
  Bertrand Thirion,
  Ga\"el Varoquaux\\
  INRIA, Parietal team, Saclay, France\\
  CEA, Neurospin, Gif-sur-Yvette, France\\
  firstname.lastname@inria.fr}

@@

To \textit{test}, on the other hand,
the neural correlates underlying mental tasks,
the general linear model (GLM) is the dominant approach \cite{friston94}.
The contribution of
individual brain voxels is estimated
according to a design matrix of experimental tasks.
Alternatively, psychophysiological interactions (PPI)
elucidate the influence of one brain region on another conditioned
by experimental tasks \cite{friston97}.
As a last example, an increasing number of
neuroimaging studies model
experimental tasks by training classification algorithms on brain signals
\cite{poldrack09decoding}.
All these methods are applied to task-associated (i.e., \textit{labeled})
data that capture brain dynamics
during stimulus-guided behavior.
Two important conclusions can be drawn.
First, the mentioned supervised neuroimaging analyses typically yield
results in a voxel space.
This ignores the fact that the BOLD
signal exhibits spatially distributed patterns
of coherent neural activity.
%
Second, existing supervised neuroimaging analyses
cannot exploit the abundance
of easily acquired resting-state data \cite{biswaldiscovery}.
These may allow better discovery of
the manifold of brain states
due to the high task-rest similarities of neural activity patterns,
as observed using ICA \cite{smith2009}
and linear correlation \cite{cole14}.
% XXX: you don't discuss the absence of information sharing across tasks / cognitions, i.e. the multi-task aspect of your work.

Both these neurobiological properties can be conjointly exploited in an
approach that is
\textit{mixed} (i.e., using rest and task data),
\textit{factored} (i.e., performing network decomposition), and
\textit{multi-task} (i.e., capitalize on neural representations shared across
mental operations).
%
The integration of brain-network discovery into
supervised classification can yield a semi-supervised learning
framework. The most relevant
neurobiological structure should hence be identified
for the prediction problem at hand.
%
Autoencoders suggest themselves because they can emulate
variants of most unsupervised learning algorithms,
including PCA, SPCA, and ICA \cite{hertz1991introduction, hinton06}.
\begin{wrapfigure}{r}{0.40\textwidth}
  \centering
  %\captionsetup{labelformat=nonumber}
    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
  \vspace{-0.7cm}
  \caption {\textbf{Model architecture}
  Linear autoencoders find an optimized compression of 79,941 brain voxels
  into $n$ unknown activity patterns by improving reconstruction from them.
  The decomposition matrix equates with the bottleneck of
  a factored logistic regression.
  Supervised multi-class learning on task data ($X_{task}$)
  can thus be guided by
  unsupervised decomposition of rest data ($X_{rest}$).
  }
\end{wrapfigure}
%
Autoencoders (AE)
are layered learning models
that condense the input data to
local and global 
representations
via reconstruction under compression prior.
%
% XXX Michael: In fact, they behave like a non-centered PCA, 
% XXX which amounts to just doing an SVD
% XXX which btw is NOT a convex optimization problem, although easy to 
% XXX calculate and with a global optimum.
% XXX Autoencoders are also not one-layered by default. Let's make sure that
% XXX we don't trip up by getting the technicalities wrong and getting 
% XXX reviewers annoyed at us because of it.
% DanilO: cool thanks. That is the stuff, we need.
They behave like a (truncated) PCA
in case of one linear hidden layer and a squared error loss
\cite{baldi1989neural}.
% This architecture yields an optimization objective
% with unique global minimum.
Autoencoders behave like a SPCA if shrinkage terms are added to the
model weights in the optimization objective.
Moreover, they have the characteristics of an ICA in case of tied weights
and adding a nonlinear convex function
at the first layer \cite{le2011ica}.
% XXX What convex transformation? Why is that relevant? Convexity is only
% XXX relevant for the loss
% Danilo: it's about adding a neural-network-like nonconvex function for
% first-layer activation
These authors further demonstrated that ICA, sparse autoencoders, and 
sparse coding are mathematically equivalent
under mild conditions.
%
Thus, autoencoders may flexibly project the neuroimaging data
onto the main directions of variation.

In the present investigation,
a linear autoencoder will be fit to
(unlabeled) rest data and
integrated as a rank-reducing bottleneck
into a multinomial
logistic regression fit to
(labeled) task data.
We can then solve the compound statistical problem of
unsupervised data representation and
supervised classification, previously studied in isolation.
%
From the perspective of dictionary learning, the first layer represents
% XXX Michael: This is the crucial difference: The first layer is an analysis
% XXX operation: it extracts information in some way. It may be extracting
% XXX activations for basis functions, in which case it would be the pseudo-
% XXX inverse of those basis functions. But unless they are orthogonal, these
% XXX weights definitely don't represent basis functions themselves. It is 
% XXX the second layer of the autoencoder that represents basis functions. If
% XXX weights are tied, they are one and the same thing, with interesting
% XXX implications. But in general, only the second layer represents a basis.
projectors to the discovered set of basis functions
which are linearly combined 
by the second layer to perform predictions \cite{olshausen96}.
% XXX Michael: ^ This needs to be changed around: This first layer obtains
% XXX activations, the second layer weights are basis functions which are
% XXX multiplied to these activations
%
Neurobiologically, this allows 
delineating a low-dimensional manifold
of brain network patterns and then 
distinguishing mental tasks
by their most discriminative linear combinations.
%
Theoretically, a reduction in model variance should be achieved by
resting-state autoencoders that
privilege the most neurobiologically
valid models in the hypothesis set.
%
Practically, neuroimaging research frequently suffers from
data scarcity. This limits the set of representations that can be
extracted from GLM analyses based on few participants.
%
We therefore contribute a computational framework that
1) analyzes many problems simultaneously
(thus finds shared representations by ``multi-task learning'')
and
2) exploits unlabeled data
(since they span a space of meaningful configurations).

\section{Methods}
%
\paragraph{Data.}
As the currently biggest openly-accessible reference dataset,
we chose resources from the Human Connectome Project (HCP)
\cite{barch2013}.
Neuroimaging task data with labels of ongoing cognitive processes
were drawn from 500
healthy HCP participants (cf. Appendix for details on datasets).
18 HCP tasks 
were selected that are known to elicit reliable neural activity
across participants (Table \ref{table_tasks}).
In sum, the HCP task data incorporated 8650 first-level activity maps
from 18 diverse paradigms administered to 498 participants (2 removed
due to incomplete data).
All maps were resampled to a common $60\times72\times60$ space of
3mm isotropic voxels and gray-matter masked (at least 10\% tissue
probability).
The supervised analyses were thus based on labeled HCP task maps with
79,941 voxels of interest representing z-values in gray matter.

@@

where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79{,}941$
voxel values from each
rest map,
$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$-dimensional hidden state
(i.e., distributed neural activity patterns), and 
$\mathbf{x' \in \mathbb{R}^{d}}$
is the reconstruction vector of the original activity map
from the hidden variables. 
% XXX use bold letters for vectors and matrices
Further, $\mathbf{W_0}$ denotes the weight matrix that
transforms
from input space into the hidden space (encoder),
$\mathbf{W_1}$ is the weight matrix for back-projection
from the hidden variables to the
output space (decoder).
$\mathbf{b_0}$ and $\mathbf{b_1}$ are corresponding bias vectors.
The model parameters $\mathbf{W_0, b_0, b_1}$ are found by
minimizing the expected squared reconstruction error
\begin{equation}
  \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
  \mathbb E\left[\| \mathbf{x} - (\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right].
\end{equation}

@@

\paragraph{Factored logistic regression.}
Our factored logistic regression model is best described as a
variant of a multinomial logistic regression.
Specifically, the weight
matrix is replaced by the product of two weight matrices with a common
latent dimension.
The later is typically much lower than the dimension of the data.
%
Alternatively, this model can be viewed as a single-hidden-layer
feed-forward neural network with a linear activation function for the
hidden layer and a softmax function on the output layer.
As the
dimension of the hidden layer is much lower than the input layer, this
architecture is sometimes referred to as a ``linear bottleneck'' in the
literature.
%
The probability of an input $\mathbf{x}$ to belong
to a class $i \in \{1, \dots, l\}$ is given by
\begin{equation}
  \begin{split}
    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_\mathcal{LR}(\mathbf x)),
  \end{split}
  \label{eq:lr}
\end{equation}
where \(f_\mathcal{LR}(\mathbf x) = \mathbf{V_1 (V_0 x + c_0) + c_1}\) computes 
multinomial logits and \(\softmax_i(x) = \exp(x_i)/\sum_j\exp(x_j)\).
The matrix $\mathbf{V_0 \in \mathbb{R}^{dxn}}$
transforms the input $\mathbf{x \in \mathbb{R}^{d}}$
into $n$ latent components
and the matrix $\mathbf{V_1 \in \mathbb{R}^{nxl}}$
projects the latent components
onto hyperplanes that reflect $l$ label probabilities.
$\mathbf{c_0}$ and $\mathbf{c_1}$ are
bias vectors.
The loss function is given by
\begin{equation}
    \mathbb E\left[{\mathcal{L_{LR}}}(\mathbf{x, y})\right] \approx %
\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1})).
\label{eq:lr_loss}
\end{equation}

\paragraph{Layer combination.}
The optimization problem of the linear autoencoder
and the factored logistic regression
are linked in two ways. First, their transformation matrices mapping from
input to the latent space are tied
\begin{eqnarray}
  \mathbf{V_0} = \mathbf{W_0}.
\end{eqnarray}
We hence search for a compression of the 79,941 voxel values into $n$ unknown
components that represent a latent code optimized for both
rest and task activity data.
Second, the objectives of the autoencoder and the factored
logistic regression are interpolated in the common loss function
% XXX Michael: In resolving conflict I opted for my change. Feel free to revert if that was intended.
\begin{equation}
{\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
+ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \Omega.
  \label{eq:loss_equ}
\end{equation}

In so doing, we search for the combined model parameters
$\theta=\{\mathbf{V_0,V_1,c_0, c_1, b_0, b_1}\}$
with respect to the (unsupervised) reconstruction error and the
(supervised) task detection.
${\mathcal{L_{AE}}}$ is devided by ${N_{X_{rest}}}$ to equilibrate both
loss terms to the same order of magnitude.
\(\Omega\) represents an ElasticNet-type regularization
that combines $\ell_1$ and $\ell_2$ penalty terms.
%$\forall p \in \theta$.

@@

\paragraph{Implementation.}
The analyses were performed in Python.
We used \textit{nilearn} to handle
the large quantities of neuroimaging data 
\cite{abrah14}
and
\textit{Theano} for automatic, numerically stable
differentiation of symbolic computation graphs
\cite{bastien2012theano, bergstra2010theano}.
All Python scripts that generated the results are
accessible online for reproducibility and reuse
% (\url{http://github.com/anonymous/anonymous}).
(\url{http://github.com/banilo/nips2015}).

@@

 \multicolumn{1}{l|}{5}   & 45.1 \%        & 32.2 \%        & 37.5 \%       & 44.2 \%        & \textbf{95.7\%}      \\
 \multicolumn{1}{l|}{20}  & 78.1 \%        & 78.2 \%        & 81.0 \%       & 63.2 \%        & \textbf{97.3\%}      \\
 \multicolumn{1}{l|}{50}  & 81.7 \%        & 84.0 \%        & 84.2 \%       & 77.0 \%        & \textbf{97.6\%}      \\
 \multicolumn{1}{l|}{100} & 81.3 \%        & 82.2 \%        & 87.3 \%       & 76.6 \%        & \textbf{97.4\%}      \\ \hline

@@

Additionally, the ``out-of-study'' performance of

@@

  \resizebox{1.00\textwidth}{!}{%
\tabcolsep=0.05cm
\begin{tabular}{l|ccccc|ccccc|ccccc}
                            & \multicolumn{5}{c|}{\textit{n = 5}}                                                                                                                                                 & \multicolumn{5}{c|}{\textit{n = 20}}                                                                                                                                                      & \multicolumn{5}{c}{\textit{n = 100}}                                                                           \\ \hline
  \textit{}                 & \multicolumn{1}{l}{$\lambda=0$}    & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0$}  & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0$      & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
  Out-of-sample\\accuracy   & \textit{6.0\%}                     & 88.9\%                             & 95.1\%                            & \textbf{96.5\%}                    & 95.7\%                           & \textit{5.5\%}                   & 97.4\%                               & \textbf{97.8\%}                     & 97.3\%                               & 97.3\%                             & \textit{6.1\%}   & 97.2\%           & 97.0\%           & \textbf{97.8\%}  & 97.4\%                             \\
  Precision (mean)          & \textit{5.9\%}                     & 87.0\%                             & 94.9\%                            & \textbf{96.3\%}                    & 95.4\%                           & \textit{5.1\%}                   & \textbf{97.4\%}                      & 97.1\%                              & 97.0\%                               & 97.0\%                             & \textit{5.9\%}   & 96.9\%           & 96.5\%           & \textbf{97.5\%}  & 96.9\%                             \\
  Recall (mean)             & \textit{5.6\%}                     & 88.3\%                             & 95.2\%                            & \textbf{96.6\%}                    & 95.7\%                           & \textit{4.6\%}                   & \textbf{97.5\%}                      & 97.5\%                              & 97.4\%                               & 97.4\%                             & \textit{7.2\%}   & 97.2\%           & 97.2\%           & \textbf{97.9\%}  & 97.4\%                             \\
  F1 score (mean)           & \textit{4.1\%}                     & 86.6\%                             & 94.9\%                            & \textbf{96.4\%}                    & 95.4\%                           & \textit{3.8\%}                   & \textbf{97.4\%}                      & 97.2\%                              & 97.1\%                               & 97.1\%                             & \textit{5.3\%}   & 97.0\%           & 96.7\%           & \textbf{97.7\%}  & 97.2\%                             \\ \hline
  Reconstr. error (norm.)   & 0.76                               & 0.85                               & 0.87                              & 1.01                               & 1.79                             & 0.64                             & 0.67                                 & 0.69                                & 0.77                                 & 1.22                               & 0.60             & 0.65             & 0.68             & 0.73             & 1.08                               \\ \hline
  Out-of-study\\accuracy    & \textit{39.4\%}                    & 60.8\%                             & 54.3\%                            & 60.7\%                             & \textbf{62.9\%}                  & \textit{77.0\%}                  & 79.7\%                               & \textbf{81.9\%}                     & 79.7\%                               & 79.4\%                             & 79.2\%           & \textbf{82.2\%}  & 81.7\%           & 81.3\%           & \textit{75.8\%}                    \\ \hline
 \caption{\textbf{Performance of SSFLogReg across model parameter choices.}
 Chance is at 5.6\%.
  \label{table_two}

We made three noteworthy observations (Table \ref{table_two}).
%
First, the most supervised estimator ($\lambda=1$) achieved in no
instance the best accuracy, precision, recall, or f1 scores on HCP data.
Classification by SSFLogReg is therefore facilitated by
imposing structure from the unlabeled rest data.
Confirmed by the normalized reconstruction error
($\mathbf{E} = \|x-\hat{x}\| / \|x\|$),
little weight on the supervised term is sufficient for good model performance
while keeping $\mathbf{E}$ low and task-map decomposition rest-like. 
%
Second, the higher the number of latent components $n$,
the higher the out-of-study performance with small values of $\lambda$.
This suggests that the presence of more rest-data-inspired hidden components
results in more effective feature representations in unrelated task data.
%
Third, for $n=20$ and $100$ (but not $5$) the purely rest-data-trained
decomposition matrix ($\lambda=0$) resulted in
noninferior out-of-study performance
of 77.0\% and 79.2\%,
respectively (Table \ref{table_two}).
This confirms that guiding model learning by task-unrelated structure
extracts features of general relevance
beyond the supervised problem at hand.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.65\textwidth]{figures/semisup_conversion.png}
  \vspace{-0.1cm}
\caption{\textbf{Weight maps of a same hidden factor ranging from
unsupervised to supervised regime}
One of the $n$ factors from the hidden layer ($\mathbf{W_0}$)
was plotted for the same data (full HCP dataset)
and
the same model choices
($n=20, \ell_1=0.1$, $\ell_2=0.1$) along
a $\lambda$-grid between
purely unsupervised ($\lambda=0.0$, \textit{top row}) and
purely supervised ($\lambda=1.0$, \textit{bottom row}) settings.
%
As qualitative evidence,
a slow transition from rest- to task-typical brain networks was observed
in brain space.
Although difficult to quantify,
rest network elements appear to get 'reassembled' to
latent factors of the LR.
This increased confidence that the improved model performance
of rest-informed fLR is not only an arbitrary
effect of spatially smooth noise.
%
All values are z-scored.
}
\label{fig_weights_conversion}
\end{figure}
